---
title:'Peer-graded Assignment: Prediction Assignment Writeup'
output:
  html_document:
    keep_md: yes
  pdf_document: default
---
# Prediction Assignment Writeup
#### Raymond Ng on 30 December 2017
#
## Overview
##### This project is to select a prediction model for predicting 20 test cases from a test dataset.
#
## Dataset Preparation
```{r}
# Load the libraries and datasets
library(caret)
library(rpart)
library(randomForest)
library(e1071)
library(gbm)
set.seed(1234)

# Set the URLs
TrainDataURL <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
TestDataURL <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"

# Download and clean the datasets
TrainingData <- read.csv(url(TrainDataURL), na.strings=c("NA","#DIV/0!",""))
TestingData <- read.csv(url(TestDataURL), na.strings=c("NA","#DIV/0!",""))

# Check for the datasets dimemsions
dim(TrainingData)
dim(TestingData)

# Delete columns with missing values
TrainingData <-TrainingData[,colSums(is.na(TrainingData)) == 0]
TestingData <-TestingData[,colSums(is.na(TestingData)) == 0]

# Delete unused columns
TrainingData <-TrainingData[,-c(1:7)]
TestingData <-TestingData[,-c(1:7)]

# Check for the datasets dimemsions
dim(TrainingData)
dim(TestingData)

```
## Dataset Partitioning

```{r}
# Partitions training dataset
PartData <- createDataPartition(TrainingData$classe, p=0.7, list=FALSE)
TrainingSet <- TrainingData[PartData, ]
TestingSet <- TrainingData[-PartData, ]
dim(TrainingSet)
dim(TestingSet)

```
#### Examine the Training Dataset by Showing its Frequency 

```{r}
plot(TrainingSet$classe, col="black", main="Frequency of different levels", xlab="classe", ylab="Frequency")

```

#### Based on the frequency distribution, it shows that each level frequency is within the same order of magnitude of each other. Level A is the most frequent, and level D is the least frequent.


## Prediction using Random Forest Model and Generalized Boosted Model

#### 1. Random Forest

```{r}
# Fitting model
model1 <- randomForest(classe ~., data=TrainingSet, method="class")

# Predicting
prediction1 <- predict(model1, TestingSet, Type="class")

# Testing
confusionMatrix(prediction1, TestingSet$classe)

```

#### 2. Generalized Boosted Model

```{r}
# Fitting model
fitControl <- trainControl(method="repeatedcv", number=5, repeats=1)
model2 <- train(classe ~., data=TrainingSet, method="gbm", trControl=fitControl, verbose=FALSE)

# Predicting
prediction2 <- predict(model2, TestingSet)

# Testing
confusionMatrix(prediction2, TestingSet$classe)

```

#### The accuracy of Random Forest is 99.64, which is higher than that by Generalized Boosted. The Random Forest
#### model will be applied to predict the 20 quiz results.

```{r}
predictTest <- predict(model1, TestingData)
predictTest

```

#### Submission

```{r}
# Create files for submission
pml_write_files = function(x){
n = length(x)
for(i in 1:n){
filename = paste0("problem_id_",i,".txt")
write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
}
}
pml_write_files(predictTest)

```
